---
title: 'Lab 1: Intro to R'
author: "Jiaxin Zheng"
date: "09/06/2024"
output:
  pdf_document: default
  openintro::lab_report: default
---

```{r load-packages, message=FALSE}
library(tidyverse)
library(openintro)
library(ggplot2)
library(dplyr)
```
### Exercise 1

```{r view-girls-counts}
arbuthnot$girls
data('arbuthnot', package='openintro')
arbuthnot

glimpse(arbuthnot)
arbuthnot$girls
```
### Exercise 2

There has decrease number in year 1640-1660. 
And huge increase number of girls baptized during 1680-1700.

```{r trend-girls}
# Insert code for Exercise 2 here
ggplot(data = arbuthnot, aes(x = year, y = girls,
                             colour= girls)) + 
  geom_point(size=3, alpha=0.5)+
  geom_smooth(method = lm,
              se=F)+
  labs(title="Number of Girls Baptized Over The Years",
     x="Year",
     y="Number of Girls")+
  theme_bw()
```

### Exercise 3

Boys baptized rate decrease.

```{r plot-prop-boys-arbuthnot}
# Insert code for Exercise 3 here
arbuthnot <- arbuthnot %>%
  mutate(boy_to_girl_ratio = boys / girls)
arbuthnot <- arbuthnot %>%
  mutate(total = boys + girls)
arbuthnot <- arbuthnot %>%
  mutate(boy_ratio = boys / total)

arbuthnot <- arbuthnot %>% 
  mutate(boy_ratio=boys/total)
  ggplot(arbuthnot, aes(x= year,
                      y= boy_ratio,
                      colour = boy_ratio))+
  geom_point(size= 3, alpha= 0.5)+
  geom_smooth(method= lm,
              se=F)+
  labs(title="Proportion of Boys Baptized Over Time",
       x="Year",
       y="Proportion of Boys")+
  theme_bw()
```
### Exercise 4

The years included in this data set is year 1940- 2002.
The data frame is num.
The variable's name are year, boys, and girls.

```{r dim-present}
# Insert code for Exercise 4 here
data('present', package='openintro')
str(present)
summary(present)
```


### Exercise 5

The number of birth records is more then number of baptism. 
The Present data set is much bigger than Arbuthnot's data

```{r count-compare}
# Insert code for Exercise 5 here
glimpse(present)
glimpse(arbuthnot)
```


### Exercise 6

Boy's birth decrease over the time. The Arbuthnot’s observation about boys being born in greater proportion than girls does not hold up in the U.S.

```{r plot-prop-boys-present}
# Insert code for Exercise 6 here
present<- present %>%
  mutate(boy_to_girl_ratio = boys / girls)
present <- present %>%
  mutate(total = boys + girls)
present <- present %>%
  mutate(boy_ratio = boys / total)

present <- present %>% 
  mutate(boy_ratio=boys/total)
  ggplot(present, aes(x= year,
                      y= boy_ratio,
                      colour = boy_ratio))+
  geom_point(size= 3, alpha= 0.5)+
  geom_smooth(method= lm,
              se=F)+
  labs(title="Proportion of Boys Over Time",
       x="Year",
       y="Proportion of Boys")+
  theme_bw()
```
### Exercise 7

Year 1961 has most total number of births in the U.S.

```{r find-max-total}
# Insert code for Exercise 7 here
present %>%
  arrange(desc(total))
```
